# Autocorrelation 

```{r}
library(ggplot2)
acfdf <- function(vec){
	vacf <- acf(vec, plot = F)
	with(vacf, data.frame(lag,acf))
}
ggacf <- function(vec){
	ac <- acfdf(vec)
	ggplot(data = ac, aes(x = lag, y = acf))+
		geom_hline(aes(yintercept = 0)) +
		geom_segment(mapping = aes(xend = lag, yend = 0))
}

tplot <- function(vec){
	df <- data.frame("X" = vec, "t" = seq_along(vec))
	ggplot(data = df, aes(x = t, y = X)) +
		geom_point() +
		geom_line()
}
```


## Independence

Two events are independent in a time series if the probability that an event at time t occurs in no way depends on the ocurrence of any event in the past or affects any event in the future. Mathematically, this is written as:

**$$\mathrm{Independence: }P \left(x_{t+1}|X_t\right) = P\left(X_{t+1}\right)$$**

## Serial Dependence /  Autocorrelation

### A definition

**If two events are independent, their corellation is 0**
That is **if $X_t$ and $X_{t+k}$ are independent, $\rho_{x_{t},x_{t+k}}=0$**

*Corollary: If the correlation between two variables is not zero, then they are not* ***independent***

In other words if $\rho_{x_{t},x_{t+k}} \neq 0$, they are ***not*** independent.


### Autocorrelation Plots

#### Dependent(ish) data
In time series we look at the **autocorrelation** between $X_t$ and $X_{t+1}$ etc (with t and t+1 it is lag 1 autocorrelation)

For example, visually, let us define a vector, $Y5$ and take its autocorrelation:

```{r, fig.show = 'hold'}
Y5 <- c(5.1,5.2,5.5,5.3,5.1,4.8,4.5,4.4,4.6,4.6,4.8,5.2,5.4,5.6,5.4,5.3,5.1,5.1,4.8,4.7,4.5,4.3,4.6,4.8,4.9,5.2,5.4,5.6,5.5,5.5)
knitr::kable(acfdf(Y5))
tplot(Y5)
ggacf(Y5)
```

#### Independent(ish) data!


Now let us look at autocorrelation of ***independent*** data

```{r, results = "hide", fig.show = "hide"}
Realization1 = gen.arma.wge(n = 250)
```


```{r, fig.show = "hold"}
tplot(Realization1)
ggacf(Realization1)
```

Now lets look at the actual value of the autocorrelation

```{r}
knitr::kable(acfdf(Realization1))
```


We see that $$ \hat{\rho}\left(t,t+1\right) = \hat{\rho}_1 = r_1 = `r mean(acfdf(Realization1)$acf)` $$
Which is more or less zero



