# Autocorrelation Concepts and Notation

```{r, include = F}
library(tswge)
library(ggplot2)
```

## Theoretical concept of $\rho$

$$\rho = \frac{E\left[\left(X-\mu_X\right)\left(Y-\mu_Y\right)\right]}{\sigma_X \sigma_Y} = \frac{\mathrm{covariance between X+y}}{\sigma_X \sigma_Y}$$

In a stationary time series, we have autocorrelation and autocovariance

## autocovariance

$$\gamma_k = E\left[\left(X_t - \mu\right)\left(X_{t+k} - \mu\right)\right]$$

## autocorrelation 

$$\rho_k =\frac{E\left[\left(X_t - \mu\right)\left(X_{t+k} - \mu\right)\right]}{\sigma_{X_t} \sigma_{X_{t+k}}} = \frac{E\left[\left(X_t - \mu\right)\left(X_{t+k} - \mu\right)\right]}{\sigma^2_X} = \frac{\gamma_k}{\sigma_X^2}$$

This works because of constant variance and constant mean!

Note that in a stationary time series:

$$\sigma_X^2 = 1E[(X_t - \mu)^2] = E[(X_t -\mu )(X_t - \mu)] = \gamma_0$$

Therefore:

$$\rho_k = \frac{\gamma_k}{\gamma_0}$$

## Stationary Covariance

Correlation is not affected by where, only by how far apart, that is:
$$ \rho_h = \mathrm{Cor} \left(X_t, X_{t+h}\right)$$

Let us try it out with this little snippet:

```{r}
ggsplitacf <-function(vec) {
  h1 <- vec[1:(length(vec)/2)]
  h2 <- vec[(length(vec)/2):length(vec)]
  first <- ggacf(vec)+ggthemes::theme_few()
  second <- ggacf(h1)+ggthemes::theme_few()
  third <- ggacf(h2)+ggthemes::theme_few()
  cowplot::plot_grid(first, second, third,
                     labels = c("original", 
                                "first half",
                                "second half"),
                     nrow = 2, align = "v")
}

```

```{r, include=F, fig.show = "hide", results='hide', echo =F}
Realize = gen.arma.wge(500,0.95,0,plot = T, sn = 784)
```



```{r, eval =F}
Realize = gen.arma.wge(500,0.95,0,plot = T, sn = 784)
```


```{r}
ggsplitacf(Realize)
```

This looks pretty stationary to me!
Let's try with some sample data:

```{r}
data("noctula")
tplot(noctula)+ggthemes::theme_few()
ggsplitacf(noctula)
```

So in this case our mean probably looks constant, our variance could be constant as well, however the ACFs are clearly different. Let's look at a different dataset:

```{r}
data(lavon)
tplot(lavon)+ggthemes::theme_few()
ggsplitacf(lavon)
```

In this case, our mean looks maybe to be constant around 495, our variance does not appear to be constant, and our ACF looks pretty good, however I would say overall this probably is not stationary, due to the variance and slightly off ACF.

## Stationary Time Series: Putting it Together

### Workflow:

* Step 1: Check condition 1
	* see if the assumption is more or less met
* Step 2: Check condition two
	* same thing
* Step 3: Check for condition three
 
If one of the things is iffy, and the other two arent, we are going to be just fine.
