# The Frequency Domain

```{r, include = F}
library(tswge)
library(tidyverse)
library(magrittr)
library(ggthemes)
```

## Review of sine and cosine


### Phase shifts
This is what a sine wave looks like

$$\delta = 0; y = \mathrm(sin)(t)$$
```{r, echo = F}
curve(sin, 0, 2*pi, xname = "t")
```

We can phase shift the sin wave by adding

RED:
$$\delta = \pi; y = \mathrm{sin}(t+\pi)$$

BLUE:
$$\delta = \frac{\pi}{2}; y = \mathrm{sin}(t+\frac{\pi}{2})$$
```{r, echo = F}
curve(sin(t + pi), 0, 2*pi, xname = "t",col = "red")
curve(sin(t + pi/2),0,2*pi,xname = "t", add = T,col="blue")
```

Now cosine

```{r, echo = F}
curve(cos, 0, 2*pi, xname = "t")
curve(cos(t + pi), 0, 2*pi, xname = "t",col = "red",add = T)
curve(cos(t + pi/2),0,2*pi,xname = "t", add = T,col="blue")
```

## Frequency and Period

Let us say $f(x)$ is a periodic function. If $f(x)$ has period $p$, then $p$ is the smallest value S.T.
$$f(x) = f(x + kp) $$

Cool

## psueudo-periodicity

Data are *psuedo-periodic* with period $p$ if $p$ is the smallest value such that a cycle appears to repeat itself.

## Aperiodicity
Aperiodic data is data in which no $p$ exists.

## Frequency

**Frequency** is calculated as : $\frac{1}{p}$

It symbolizes the number of cycles per unit(number of periods per unit)

## Frequency Terminology
Frequency of a trigonometruc function:

Let $f(x)=\sin\left( 2\pi f t \right)$

Then we have that :

$$p = 1/f$$ $$\nu = f$$

Consider:
$$\sin(Bt + C)$$
Then 
$$p = \frac{2 \pi}{B}$$
$$\nu = \frac{B} {2 \pi}$$

## Composite sine function

```{r, code = readLines('R/compositesine.R')}
```

this is a signal with three frequencies!
