# ARIMA and ARUMA
Ok this is gonna be a light week, basically, in our notation, we have that ARIMA is an ARMA model with a wandering term, eg an equation with a root at exactly 1, and just one of them, while an ARUMA is a model where the backshift operator is squared. Mathematically we have:

## ARIMA
$$(1 - B)^n X_t = a_t$$
is an ARIMA (0,n,0) model.

## ARUMA
$$(1 - B^n) X_t = a_t $$ 
is an ARUMA(0,n,0) model, or an ARIMA(0,0,0) model with a seasonal component of order n. This is pretty easy.

## But what do they do????
ARIMA adds a wandering component, since the root is exactly one, while ARUMA is a seasonal model, because of sines and cosines and Euler's theorem etc. Dont worry too much about it, try some stuff out with `artrans.wge, gen.arima.wge, gen.aruma.wge` etc. Here are some wrapper functions for artrans.wge:

### ARIMA Transform

```{r, results = "hide"}
library(tswge)
arimatr <- function(xs,d){
	f <- artrans.wge(xs, phi.tr = 1)
	if (d == 1) {
		f
	}
	else {
		arimatr(f,d-1)
	}
}

x = gen.arima.wge(500, phi = c(.6,-0.8), d = 2, theta = .3, sn = 35, plot = F)
d <- artrans.wge(x, phi.tr = 1)
d2 <- artrans.wge(d, phi.tr = 1)
d3 <- arimatr(x,2)
```

### ARUMA Transform

```{r, results = "hide"}
arumatr <- function(xs,s){
	artrans.wge(xs, phi.tr = c(rep(0,s-1),1))
}

x2 <- gen.aruma.wge(n = 200, s=7, plot = F)
arumatr(x2,7)
```

Again by transforming away the non stationary nonsense we end up with pure white noise in this case. We can check that with aic5.wge, it should be very confused
